### 2017-02-17 CS798-ConvexityAndOptimization/assignments/assignment_01.tex
p7

%Q6
\section{}
%Pa
\subsection{a}
\paragraph{}
Let $\cE_1:= \{P_1 u + q_1 : ||u||_2 \leq 1\}$ and $\cE_2:= \{P_2 u + q_2 : ||u||_2 \leq 1\}$ be given ellipsoids. We want to find a separating hyperplane $(a,b)$ for these ellipsoids, that is:
\begin{align*}
a^Tx + b &\geq 0 &\forall x \in \cE_1 \\
a^Tx+b &\leq 0 &\forall x \in \cE_2.
\end{align*}
This can be formulated as an optimization feasibility problem as
\begin{align*}
\min\ 0\\
\text{s.t.}\ a^Tx + b &\geq 0 &\forall x \in \cE_1\\
a^Tx + b &\leq 0 &\forall x \in \cE_2.
\end{align*}
This problem is equivalent to
\begin{align*}
\min\ 0\\
\text{s.t.}\ a^T(-x)&\leq b &\forall x \in \cE_1\\
a^Tx &\leq -b &\forall x \in \cE_2.
\end{align*}
Which is equivalent to
\begin{align*}
\min\ 0\\
\text{s.t.}\  
\sup\{ a^T(-x) : x \in \cE_1\}&\leq b\\
\sup\{a^Tx : x\in \cE_2\} &\leq -b.
\end{align*}
Equivalently
\begin{align*}
\min\ 0\\
\text{s.t.}\  
\sup\{ a^T(-P_1 u - q_1) : ||u||_1 \leq 1\}&\leq b\\
\sup\{a^T(P_2 u + q_2) : ||u||_2 \leq 1\} &\leq -b.
\end{align*}
Which is equivalent to
\begin{align*}
\min\ 0\\
\text{s.t.}\  
a^T(-q_1) + ||P_1a||_2 &\leq b\\
a^Tq_2 + ||P_2a||_2 &\leq -b.
\end{align*}
Setting $A_1 = P_1$, $b_1 = 0$, $c_1=q_1$, $d_1 = b$, and $A_2 = P_2$, $b_2 = 0$, $c_2 = -q_2$, and $d_2 = -b$ we obtain the exact form of a Second Order Conic Program as described in the general formulation of such problems. $\blacksquare$
\subsection{b}
\paragraph{}
We will write the Second-Order Cone Program (SOCP):
\begin{align*}
\min &f^Tx \\
\text{s.t.} ||A_ix + b_i||_2 &\leq c^T_i x + d_i, &1\leq i \leq m \\
Fx &= g,
\end{align*}
as a Semidefinite Program (SDP) in inequality form (see section $4.6.2$ of Boyd and Vandenberghe) :
\begin{align*}
\min\ &f^Tx \\
\text{s.t.}\ G + \sum_{i=1}^n x_i F_i &\preccurlyeq 0 \\
Fx &=g.
\end{align*}
Observe that it is enough to show that a constraint of the form $||Ax + b|| \leq c^Tx + d$ can be written as $G + \sum_{i=1}^n x_i F_i \preccurlyeq 0$ as then we could write each constraint $||A_ix + b_i||_2 \leq c^T_i x + d_i$ for $i\in [m]$ as $G^i + \sum_{j=1} x_j F^i_j$, and stack them as blocks as
$$G = \begin{bmatrix} G^1 & & \\ & \ddots & \\ & & G^m \end{bmatrix} \quad F_i = \begin{bmatrix} F^1_i & & \\ & \ddots & \\ & & F^m_i \end{bmatrix}$$
to obtain an equivalent single semidefinite constraint in place of $m$ semidefinite constraints.
\paragraph{Lemma 6b}
We will need the following lemma using Schur's complement: For any $x\in \R^n$ and $c\in \R$ we have $||x||_2 \leq c$ if and only if
$$X = \begin{bmatrix} c &x^T \\ x & cI\end{bmatrix} \succcurlyeq 0.$$
\paragraph{Proof of Lemma 6b}
If $c = 0$ this is trivial, so we may assume that $c\neq 0$. In this case we invoke Schur's Complement Lemma and thus
$$X \succcurlyeq 0$$
if and only if $c \geq 0$ and $tI - \frac{1}{c}xx^T \succcurlyeq 0$. Which holds if and only if $c^2I - xx^T \succcurlyeq 0$. Now the eigenvalues of $c^2I - xx^T$ are non-negative if and only if $c^2 - ||x||^2_2 \geq 0$ (since $c^2$ is the least eigenvalue of $c^2I$ and $||x^2_2$ is the largest eigenvalue of $xx^T$), which holds if and only if $||x||_2 \leq c$ as desired. $\blacksquare$
\paragraph{}
Now applying Lemma $6b$ to our SOCP cone constraint we have $||Ax +b||_2 \leq c^Tx + d$ (suppose $x\in\R^n$ and $A \in \R^{r\times n}$) if and only if
$$\begin{bmatrix} c^Tx + d & (Ax+b)^T \\ Ax + b &(c^T+d)xI \end{bmatrix} \succcurlyeq 0.$$
Let $a^1, \dots, a^r$ be the rows of $A$. Separating out we obtain:
$$\begin{bmatrix} d & b^T \\ b & dI \end{bmatrix} + \sum_{i=1}^n x_i \begin{bmatrix} c_i & a^1_i & \dots & a^r_i \\ a^1_i \\ \vdots & &c_iI\\ a^r_i \end{bmatrix} \succcurlyeq 0.$$
Thus $$G = \begin{bmatrix} d & b^T \\ b & dI \end{bmatrix}$$
and 
$$F_i = \begin{bmatrix} c_i & a^1_i & \dots & a^r_i \\ a^1_i \\ \vdots & &c_iI\\ a^r_i \end{bmatrix}$$
we have
$$G + \sum_{i=1}^n F_i \succcurlyeq 0$$
if and only if
$$||Ax+b||_2 \leq c^Tx + d.$$
Therefore we have shown that any SOCP can be written as an SDP. $\blacksquare$
\end{document}

>>>>

%Q6
\section{}
%Pa
\subsection{a}
\paragraph{}
Let $\cE_1:= \{P_1 u + q_1 : ||u||_2 \leq 1\}$ and $\cE_2:= \{P_2 u + q_2 : ||u||_2 \leq 1\}$ be given ellipsoids. We want to find a separating hyperplane $(a,b)$ for these ellipsoids, that is:
\begin{align*}
a^Tx + b &\geq 0 &\forall x \in \cE_1 \\
a^Tx+b &\leq 0 &\forall x \in \cE_2.
\end{align*}
This can be formulated as an optimization feasibility problem as
\begin{align*}
\min\ 0\\
\text{s.t.}\ a^Tx + b &\geq 0 &\forall x \in \cE_1\\
a^Tx + b &\leq 0 &\forall x \in \cE_2.
\end{align*}
This problem is equivalent to
\begin{align*}
\min\ 0\\
\text{s.t.}\ a^T(-x)&\leq b &\forall x \in \cE_1\\
a^Tx &\leq -b &\forall x \in \cE_2.
\end{align*}
Which is equivalent to
\begin{align*}
\min\ 0\\
\text{s.t.}\  
\sup\{ a^T(-x) : x \in \cE_1\}&\leq b\\
\sup\{a^Tx : x\in \cE_2\} &\leq -b.
\end{align*}
Equivalently
\begin{align*}
\min\ 0\\
\text{s.t.}\  
\sup\{ a^T(-P_1 u - q_1) : ||u||_1 \leq 1\}&\leq b\\
\sup\{a^T(P_2 u + q_2) : ||u||_2 \leq 1\} &\leq -b.
\end{align*}
Which is equivalent to
\begin{align*}
\min\ 0\\
\text{s.t.}\  
a^T(-q_1) + ||P_1a||_2 &\leq b\\
a^Tq_2 + ||P_2a||_2 &\leq -b.
\end{align*}
Setting $A_1 = P_1$, $b_1 = 0$, $c_1=q_1$, $d_1 = b$, and $A_2 = P_2$, $b_2 = 0$, $c_2 = -q_2$, and $d_2 = -b$ we obtain the exact form of a Second Order Conic Program as described in the general formulation of such problems. $\blacksquare$
\subsection{b}
\paragraph{}
We will write the Second-Order Cone Program (SOCP):
\begin{align*}
\min &f^Tx \\
\text{s.t.} ||A_ix + b_i||_2 &\leq c^T_i x + d_i, &1\leq i \leq m \\
Fx &= g,
\end{align*}
as a Semidefinite Program (SDP) in inequality form (see section $4.6.2$ of Boyd and Vandenberghe) :
\begin{align*}
\min\ &f^Tx \\
\text{s.t.}\ G + \sum_{i=1}^n x_i F_i &\preccurlyeq 0 \\
Fx &=g.
\end{align*}
Observe that it is enough to show that a constraint of the form $||Ax + b|| \leq c^Tx + d$ can be written as $G + \sum_{i=1}^n x_i F_i \preccurlyeq 0$ as then we could write each constraint $||A_ix + b_i||_2 \leq c^T_i x + d_i$ for $i\in [m]$ as $G^i + \sum_{j=1} x_j F^i_j$, and stack them as blocks as
$$G = \begin{bmatrix} G^1 & & \\ & \ddots & \\ & & G^m \end{bmatrix} \quad F_i = \begin{bmatrix} F^1_i & & \\ & \ddots & \\ & & F^m_i \end{bmatrix}$$
to obtain an equivalent single semidefinite constraint in place of $m$ semidefinite constraints.
\paragraph{Lemma 6b}
We will need the following lemma using Schur's complement: For any $x\in \R^n$ and $c\in \R$ we have $||x||_2 \leq c$ if and only if
$$X = \begin{bmatrix} c &x^T \\ x & cI\end{bmatrix} \succcurlyeq 0.$$
\paragraph{Proof of Lemma 6b}
If $c = 0$ this is trivial, so we may assume that $c\neq 0$. In this case we invoke Schur's Complement Lemma and thus
$$X \succcurlyeq 0$$
if and only if $c \geq 0$ and $tI - \frac{1}{c}xx^T \succcurlyeq 0$. Which holds if and only if $c^2I - xx^T \succcurlyeq 0$. Now the eigenvalues of $c^2I - xx^T$ are non-negative if and only if $c^2 - ||x||^2_2 \geq 0$ (since $c^2$ is the least eigenvalue of $c^2I$ and $||x^2_2$ is the largest eigenvalue of $xx^T$), which holds if and only if $||x||_2 \leq c$ as desired. $\blacksquare$
\paragraph{}
Now applying Lemma $6b$ to our SOCP cone constraint we have $||Ax +b||_2 \leq c^Tx + d$ (suppose $x\in\R^n$ and $A \in \R^{r\times n}$) if and only if
$$\begin{bmatrix} c^Tx + d & (Ax+b)^T \\ Ax + b &(c^T+d)xI \end{bmatrix} \succcurlyeq 0.$$
Let $a^1, \dots, a^r$ be the rows of $A$. Separating out we obtain:
$$\begin{bmatrix} d & b^T \\ b & dI \end{bmatrix} + \sum_{i=1}^n x_i \begin{bmatrix} c_i & a^1_i & \dots & a^r_i \\ a^1_i \\ \vdots & &c_iI\\ a^r_i \end{bmatrix} \succcurlyeq 0.$$
Thus setting $$G = -\begin{bmatrix} d & b^T \\ b & dI \end{bmatrix}$$
and 
$$F_i = -\begin{bmatrix} c_i & a^1_i & \dots & a^r_i \\ a^1_i \\ \vdots & &c_iI\\ a^r_i \end{bmatrix}$$
we have
$$G + \sum_{i=1}^n F_i \preccurlyeq 0$$
if and only if
$$||Ax+b||_2 \leq c^Tx + d.$$
Therefore we have shown that any SOCP can be written as an SDP. $\blacksquare$

%Q7
\section{}
\paragraph{}
Consider the following optimization problem $(P)$:
\begin{align*}
\min\ &e^{-x} \\
\text{s.t.}\ \frac{x^2}{y} &\leq 0,
\end{align*}
with domain $\cD = \{(x,y) : y > 0 \}$. The Lagrangian function $g(\lambda)$ is
$$ g(\lambda) := \inf_{x,y} \{ e^{-x} + \lambda \frac{x^2}{y}\}.$$
We will study the behaviour of $g$ to understand the value of the dual problem.
\paragraph{}
The dual problem $(D)$ is:
\begin{align*}
\max\ &g(\lambda) \\
\lambda &\geq 0.
\end{align*}
We consider two cases: when $\lambda = 0$ and when $\lambda > 0$.
\paragraph{}
When $\lambda = 0$, 
$$g(\lambda) = \inf_x e^{-x} = 0.$$
Now when $\lambda > 0$ observe that $y = \lambda x^4 > 0$ for $x > 0$ and so
$$\lim_{x \rightarrow \infty, y=x^4} (e^{-x} + \lambda\frac{x^2}{y}) = \lim_{x\rightarrow \infty} (e^{-x} + \lambda\frac{x^2}{\lambda x^4}) = \lim_{x\rightarrow \infty}( e^{-x} + \frac{1}{x^2}) = 0.$$
Since $g(\lambda) \geq 0$ this implies $g(\lambda) = 0$. Therefore in either case $g(\lambda) = 0$. Thus $(D)$ is equal to:
\begin{align*}
\max\ &0 \\
\lambda &\geq 0.
\end{align*}
And thus if we let$d^*$ be the optimal value of $(D)$, then $d^*=0$.
\paragraph{}
Now we return our attention to $(P)$. Since $x^2 \geq 0$ for all $x$ and $y > 0$ we observe that
$$\frac{x^2}{y} \leq 0 \iff x = 0.$$
Thus all feasible $(x,y)$ for $(P)$ have objective value $e^{-0} = 1$. So if we let $p^*$ be the optimal value of $(P)$ then $p^* = 1$.
\paragraph{}
Therefore $p^* = 1 > 0 = d^*$, so strong duality does not hold for this problem. $\blacksquare$
\end{document}
